### 2013-11-12 pldi2014/content/abstract.tex
Factoring out.

\begin{abstract}
Dynamic contract enforcement costs time and space to perform, sometimes dominating the execution cost of programs.
Gathering the static information necessary to discharge them before program execution may shift the perspective of contracts to a lower level and force analyses to contend with the expediencies of implementation representations.
In systems where contracts can be attached dynamically, as well as enforced, analyses must consider any influence a given contract has on program behavior as well as that of any contract attached or enforced by it, directly or indirectly.
We present a static analysis which can extract and reason about these relationships to safely discharge contracts when they are transparent to the program.
We analyze Racket to demonstrate the ability of the analysis to cope with dynamic contract dependencies and particular implementation constraints.
Our analysis integrates the reasoning of dependency, blame, and termination which allows us to considerably reduce run-time contract prevalence.

%Dynamic contract checks cost time and space to perform.
%In languages where contracts may be attached dynamically, as well as checked, costs compound until they dominate actual program execution.
%Gathering the static information necessary to discharge them \emph{before} program execution may shift the view of contracts to a lower level and force analyses to contend with implementation representations.
%In languages where contracts may be attached dynamically, as well as checked, judging whether a contract can be safely discharged requires establishing that \emph{every} contract whose attachment or checking 
%Dynamic contract checks---especially for higher-order values---take time and space to perform.
%In practice, the cost of these checks can dominate the run-time cost of the executing the program.
%It is therefore desirable to discharge them statically where possible.
%However, implementation realities may mandate a shift in perspective as the view of contracts at low levels may vary.
%is not as simple as determining when they directly influence the meaning of the program
%underscores the reality that, in correct programs, contracts should be transparent
%In systems where contracts may be attached to higher-order values dynamically and at the hand of another contracted value, the scope 
%If 
%We focus on demodularized Racket programs in which dynamic procedure contracts are expressed in terms of higher-order interposition and present an analysis for whole programs to statically detect and discharge benign contracts.
%Utilizing a combination of introspective pushdown control-flow analysis and termination analysis, we demonstrate cases in which our analysis significantly decreases the run-time burden of specification enforcement.
\end{abstract}