### 2013-11-12 pldi2014/content/limitations.tex

\section{Limitations}

Our calculus has an impoverished value domain, so our termination judgments of guards are nearly complete in practice as a result of only a simple flow analysis.
(Because, in practice, guards are not recursive themselves or any procedures in the call graph. Essentially, call graphs for guard procedures are trees.)
Contracts on compound data complicate matters because the data may be defined inductively which makes recursive checking natural.
We can accomodate this in particular by integrating size-change analysis~\cite{lee2001size} into our abstract interpretation.
This would be novel as higher-order size-change analysis~\cite{sereni2007termination} has heretofore been applied only to variants of $k$-CFA~\cite{shivers1991control}.

\scheme{app-values} increases the expressiveness of the language.
(We do not have to specify the number of arguments to a function statically.)
The structure of a list is recursive.
We can untie the recursive definition by directing through the store; we took this approach with environments.
We could look into this, but that means the length of the list could vary.
That could \emph{really} destroy precision, but we need to investigate it further.

\emph{rest values} increase the expressiveness of the language. They are very related to \scheme{app-values}.
\scheme{chap-op} as a primitive (as opposed to a form) increases the expressiveness of the language.

Our analysis applies equally well if we extend the language with more referentially transparent primitives or macro-expressible constructs such as \scheme{and} and \scheme{or} forms.
We can even add an error-catching facility but we must be careful in our analysis.

In Racket, the \scheme{chaperone-of?} relation is accessible through a predicate procedure.
However, its presence in Racket has different implications than its introduction to \chapcalc, which can use it to determine whether two procedures are intensionally equivalent.
(In Racket, its use in that fashion is restricted to determining whether two references indicate the same procedure.)

Chaperones are intended to be treated like the original value.
% box?, procedure?, etc. will respond the same way
% they evaluate equal? to the values they wrap, but not eq?
Procedure chaperones are considered to have the same arity of the procedures they wrap.
The \scheme{chaperone?} predicate is designated as the single way to detect chaperones at runtime, though its use is discouraged.

\subsection{Handling \emph{handle}}

Exception handlers prevent certain errors from escaping a particular dynamic extent.
Consider the program
\begin{schemedisplay}
(handle
  (chap-op + (raise) (raise))
  -)
\end{schemedisplay}
wherein an error is always raised and the handler always triggered.
Even though an exception never escapes the dynamic extent of the \scheme{handle} expression, replacing the \scheme{chap-op} application with its first argument results in a different program. 
This means that we must consider whether a \scheme{chap-op} application ever raises an error, and not simply if that error escapes a delimited extent.
% being able to catch errors means that errors can change the meaning of a program arbitrarily--even if it results in a value.
% this means that we need to preserve errors
% chaperone-of? within the language can change the behavior, right? if it was a chaperone before, removing the chaperone makes it still chaperone-of? it's chaperone? that makes it possible to detect changes, but that is discouraged, and we will leave it out of our language.

chaperone and impersonator properties