### 2013-11-12 pldi2014/paper.tex

\documentclass{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{slatex}

\setspecialsymbol{->}{$\rightarrow$}
\setspecialsymbol{LAMBDA}{$\lambda$}
\setspecialsymbol{GE}{$\ge$}
\setspecialsymbol{EPS}{$\epsilon$}

\setkeyword{define/contract}

\mathchardef\mhyphen="2D
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{PLDI '14}{Month d--d, 2014, Edinburgh, Scotland} 
\copyrightyear{2014} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Static Discharge of Dynamic, Higher-Order Contracts}
%\subtitle{Subtitle Text, if any}

%\authorinfo{Kimball Germane\and Matthew Might}
%           {University of Utah}

\authorinfo{name(s) withheld}
           {}

\begin{document}

\newcommand{\chapcalc}[0]{\ensuremath{\lambda_{\mathcal{C}}}}

% TODO:
% talk about needing to keep track of every frame of all functions which haven't returned but we're only interested in the chaperone pos and neg frames because they tell us the active chaperones--we can (over-)approximate the active chaperones with the flow analysis.
% make the "dynamic blame" be an actual transition to a blame state or something like that--no matter the error, we want to look at the dependent, so maybe we have no dynamic blame and just have an "empty" blame knowing that the implementation can determine blame actually but we don't need to pretend (because we need all the active, not the youngest active)
% format everything
% revisit factorial at the end
% visit church encoded factorial
% visit the impact of domains
% motivate need for termination analysis
% figure out a way to present it as product and not sum of contributions
% take out church factorial fluff
% write related work section
% make it clear that we can prevent the DYNAMIC installation of contracts

\maketitle

\input{content/abstract.tex}
\input{content/introduction.tex}
\input{content/interposition.tex}
\input{content/calculus.tex}
\input{content/discharge.tex}
\input{content/analysis.tex}
\input{content/limitations.tex}
\input{content/related.tex}
\input{content/conclusion.tex}

% behavioral contracts
% monitoring
% projection
% discharge
% demotion or promotion
% compliance
% chaperone vs. chaperoned operator
%

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

%\acks

%These are the acks.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{paper}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices