### 2013-11-15 pldi2014/content/calculus.tex
Formatting semantics.

\section{Chaperone Calculus}

In order to explicate a static analysis of higher-order interposition, we define \chapcalc, or \emph{chaperone calculus}, an extension of the A-normalized $\lambda$-calculus with interposition facilities.

As only a moderate extension of the $\lambda$-calculus, \chapcalc\ lacks many Racket features.
%Mutation, in particular, is conspicuously absent.
We discuss the integration of other language features in Section \ref{sec:limitations}.

\subsection{\chapcalc\ Syntax}

The syntax of \chapcalc\ can be seen in Figure \ref{fig:syntax}.

\chapcalc\ inherits application, abstraction, and variable terms from the $\lambda$-calculus.
It includes simple values from the domain of integers and booleans as well as a set of first-class primitive procedures which operate over these values.
The inclusion of primitive procedures and introduction of impersonators and chaperones enriches the class of operators formerly inhabited only by closures.
This enrichment is masked somewhat in the language behind a single discriminator---\scheme{operator?}---for any members of the class.

The syntactic class \ae\ signifies \emph{atomic expressions} and includes $\lambda$-terms, variables, and the simple values of booleans, integers, and primitive procedures, which include \scheme{imp-op} and \scheme{chap-op}.
Expressions classified as such will never produce an error or diverge.

The syntactic class $c$ denotes procedure application sites which are formed by an atomic operator expression followed by a sequence of atomic operand expressions.

The syntatic class of expressions, denoted $e$, includes \scheme{let}, and \scheme{if} expressions.
Concrete syntax in examples may include \scheme{letrec} which is defined simply in terms of \scheme{let} and the \scheme{Y} combinator, as well as \scheme{and} and \scheme{or} defined in terms of \scheme{if}.

Every term is annotated with a unique label $\ell$ to keep otherwise identical terms distinguishable, but labels will be made explicit only for application sites.

A program is a term in the grammar with no free variables and in which all identifiers in the same binding form are pairwise distinct.

\newcommand{\vx}[0]{\mathbf{x}}

\newcommand{\appe}[2]{(#1\,#2)^\ell}
\newcommand{\lame}[2]{(\lambda\,(#1)\,#2)}
\newcommand{\chae}[2]{(\mathit{chaperone\mhyphen operator}\,#1\,#2)^\ell}
\newcommand{\impe}[2]{(\mathit{impersonate\mhyphen operator}\,#1\,#2)}
\newcommand{\lete}[3]{(\mathbf{let}\,((#1)\,#2)\,#3)}
\newcommand{\ife}[3]{(\mathbf{if}\,#1\,#2\,#3)}

\newcommand{\true}[0]{\mathrm{\#t}}
\newcommand{\false}[0]{\mathrm{\#f}}

\newcommand{\stxclass}[4]{$#1\in\mathbf{#2}$ &::=& #3 & #4}
\newcommand{\stxclasscont}[2]{&$|$& #1 & #2}

\setspecialsymbol{ae}{$\ae$}

\setlength{\tabcolsep}{1pt}
\begin{figure}
\label{fig:syntax}

\begin{tabular}{ r r l l }
\stxclass{e}{Exp}{\scheme|(let ((x ...) c) e)|}{let}\\
\stxclasscont{\scheme|(if ae e e)|}{if}\\
\stxclasscont{$c$}{}\\
\stxclasscont{\ae}{}\\
\stxclass{c}{Call}{\scheme|(ae ae ...)|$^\ell$}{application}\\
\stxclass{\ae}{AExp}{$\mathit{lam}$ $|$ $p$ $|$ $x$ $|$ $n$ $|$ $t$}{atomic expressions}\\
\stxclass{\mathit{lam}}{Lam}{\scheme|(LAMBDA (x ...) e)|}{lambda terms}\\
\stxclass{p}{Prim}{\scheme|imp-op| $|$ \scheme|chap-op| $|$ \scheme|values| $|$ \scheme|raise|}{primitives}\\
\stxclasscont{\scheme|<| $|$ \scheme|>| $|$ \scheme|=| $|$ \scheme|+| $|$ \scheme|-| $|$ \scheme|not|}{}\\
\stxclasscont{\scheme|operator?| $|$ \scheme|integer?| $|$ \scheme|boolean?|}{}\\
\stxclass{x}{Var}{an infinite set of variables}{}\\
\stxclass{n}{Int}{$0\,|\,1\,|\,-1\,|\,2\,|\,-2\,|\,\dots$}{integers}\\
\stxclass{t}{Bool}{$\true\,|\,\false$}{booleans}\\
\stxclass{\ell}{Lab}{an infinite set of labels}{}
\end{tabular}

\caption{The syntax of \chapcalc.}
\end{figure}

% results are:
% values: closures, chaperones, impersonators, primitives, integers, booleans
% errors: error, blame L, blame +L, blame -L

% (e e1 ... en),\rho,\sigma,\kappas
% v = A(e,\rho,\sigma), vi = A(ei,\rho,\sigma)
% case v
%   ((lambda s e'),\rho')
%   if s compatible with n
%     (e',\rho'',\sigma',\kappas) where \rho'' and \sigma' bind vis according to s
%     error
%   primitive
%   if primitive arity compatible with n
%     \delta(primitive,v1,...,vn)
%     error
%   (chaperone L f neg pos)
%   (lambda vs
%     (let ([vs' (app-values neg vs)])
%       (if (= (length vs)
%              (length vs'))
%           (if (and-map chaperone-of? vs' vs)
%               (let ([rs' (app-values f vs')])
%                 (if (arity-comp? pos rs')
%                     (let ([rs (app-values pos rs')])
%                       (if (= (length rs')
%                              (length rs))
%                           (if (and-map chaperone-of? rs rs')
%                               (app-values values rs)
%                               (error)) (blame + guard)
%                           (error))) (blame + guard)
%                     (error))) (blame + guard)
%               (error)) (blame - guard)
%           (error)))) (blame - guard)

\newcommand{\A}[3]{\mathcal{A}(#1,#2,#3)}
\newcommand{\arity}[1]{\ensuremath{\mathrm{arity}(#1)}}

\newcommand{\dynbla}[1]{\mathrm{dynamic\mhyphen blame}(#1)}

\newcommand{\nttwo}[3]{\ensuremath{\mathbf{#1}(#2,#3)}}
\newcommand{\ntthr}[4]{\ensuremath{\mathbf{#1}(#2,#3,#4)}}
\newcommand{\ntfou}[5]{\ensuremath{\mathbf{#1}(#2,#3,#4,#5)}}

\newcommand{\ks}[0]{\gamma^*}
\newcommand{\vv}[0]{\mathbf{v}}
\newcommand{\vvp}[0]{\mathbf{v'}}
\newcommand{\sexp}[4]{\mathbf{eval}(#1,#2,#3,#4)}
\newcommand{\sval}[3]{\mathbf{value}(#1,#2,#3)}
\newcommand{\sapp}[4]{\mathbf{apply}(#1,#2,\ell,#3,#4)}
\newcommand{\simpz}[6]{\mathbf{imp_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\simpo}[5]{\mathbf{imp_1}(#1,#2,#3,#4,#5)}
\newcommand{\schaz}[6]{\mathbf{chap_0}(#1,#2,#3,#4,#5,#6)}
\newcommand{\schao}[5]{\mathbf{chap_1}(#1,#2,#3,#4,#5)}
\newcommand{\serr}[3]{\mathbf{error}(#1,#2,#3)}
\newcommand{\sbla}[2]{\mathrm{blame}_{#1}(#2)}
\newcommand{\app}[4]{\mathrm{apply}(#1,#2,#3,#4)}
\newcommand{\reda}[2]{\begin{equation}\begin{aligned}& #1\\\rr\, & #2\end{aligned}\end{equation}}
\newcommand{\red}[2]{$#1\rr#2$}
\newcommand{\redtwo}[5]{\begin{math}#1\rr\begin{cases}#2 & #3\\ #4 & #5\end{cases}\end{math}}
\newcommand{\redthr}[7]{\begin{math}#1\rr\begin{cases}#2 & #3\\ #4 & #5\\#6 & #7\end{cases}\end{math}}

\newcommand{\clo}[2]{\nttwo{clos}{#1}{#2}}
\newcommand{\imp}[2]{\ntthr{imp}{\ell}{#1}{#2}}
\newcommand{\cha}[2]{\ntthr{chap}{\ell}{#1}{#2}}

\newcommand{\letk}[3]{\ntthr{let}{#1}{#2}{#3}}

\newcommand{\impcwk}[2]{\ntthr{imp\mhyphen neg}{\ell}{#1}{#2}}
\newcommand{\impcfk}[1]{\nttwo{imp\mhyphen fun}{\ell}{#1}}
\newcommand{\impcrk}[1]{\nttwo{imp\mhyphen pos}{\ell}{#1}}

\newcommand{\chacwk}[2]{\ntthr{chap\mhyphen neg}{\ell}{#1}{#2}}
\newcommand{\chacfk}[1]{\nttwo{chap\mhyphen fun}{\ell}{#1}}
\newcommand{\chacrk}[1]{\nttwo{chap\mhyphen pos}{\ell}{#1}}

\newcommand{\blame}[1]{\ensuremath{\mathrm{blame}(#1)}}
\newcommand{\impblame}[1]{\ensuremath{\mathrm{imp\mhyphen blame}(#1)}}
\newcommand{\chablame}[1]{\ensuremath{\mathrm{cha\mhyphen blame}(#1)}}
\newcommand{\noblame}[0]{\ensuremath{\mathrm{noblame}}}

\newcommand{\rr}{\longrightarrow}
\newcommand{\rrs}{\longrightarrow^{*}}

\subsection{$\chapcalc$ Semantics}

We define the semantics of \chapcalc\ in terms of an abstract machine similar to Felleisen's CESK machine~\cite{felleisen1987calculus} to facilitate the eventual pushdown analysis.
Figure~\ref{fig:states} presents the definition of the abstract machine space.

We use boldface to denote a vector of objects in that category; e.g., $\vv$ denotes a vector $\langle v_1,\dots,v_n\rangle$ for some natural number $n$.
We use a star ($^*$) for the same purpose when boldface is not available; e.g., $\gamma^*$ denotes a vector $\langle\gamma_1,\dots,\gamma_n\rangle$.
The notation $x::\vx$ denotes the prepend of $x$ to the vector $\vx$.
(If $\vx=\langle x_1,x_2,\dots,x_n\rangle$, then $x::\vx=\langle x,x_1,x_2,\dots,x_n\rangle$.)

First we define a set of metafunctions that the semantics relies on.

\subsubsection{Program Injection}

A program is injected into a machine state with the function $\mathcal{I} : e\rightarrow\mathbf{State}$ defined as $\mathcal{I}(p)=\sexp{\langle\rangle}{\perp}{\perp}{p}$.

\subsubsection{Atomic Expression Evaluation}

\newcommand{\Aeval}[1]{\ensuremath{\mathcal{A}(\sigma,\rho,#1)}}

The metafunction $\mathcal{A}$ is used to evaluate atomic expressions and is defined by
\begin{tabular}{ r l l }
\Aeval{\mathit{lam}} &= \clo{\mathit{lam}}{\rho} & $\lambda$-terms\\
\Aeval{x}            &= $\sigma(\rho(x))$ & variables\\
\Aeval{n}            &= $n$ & integers\\
\Aeval{t}            &= $t$ & booleans\\
\Aeval{p}            &= $p$ & primitives\\
\end{tabular}

\subsubsection{Variable Binding}

\newcommand{\bind}[2]{\ensuremath{\mathrm{bind}(\sigma,\rho,#1,#2)}}
\newcommand{\alloc}[1]{\ensuremath{\mathrm{alloc}(#1)}}

Rather than an environment binding a variable to a value, it binds a variable to an address which is bound to a value by the store.
(This indirection will make later semantic changes more subdued.)
In consequence, the a new variable-value binding touches both the environment and the store.

We define $\bind{x}{v}=(\sigma',\rho')$ where $\sigma'=\sigma[a\rightarrow v]$, $\rho'=\rho[x\rightarrow a]$, and $a=\alloc{x}$ where $\alloc{x}$ simply returns a fresh address for $x$.
We also fold $\mathrm{bind}$ over identically-sized variable vectors $\vx$ and value vectors $\vv$.

\subsubsection{chaperone-of?}

\newcommand{\chapof}[2]{\ensuremath{\mathrm{chaperone\mhyphen of?}(#1,#2)}}

The relation \chapof{v}{v'} holds when parts of $v$ can be derived from corresponding parts of $v'$ through \scheme{chap-op}.
We lift $\mathrm{chaperone\mhyphen of?}$ element-wise over vectors.

%It holds if any of the following hold.
%(This definition is not given in terms of sequential pattern matching. Two chaperones satisfy the relation if they meet the conditions of the first clause \emph{or} the second clause.)

\begin{itemize}
\item[] \chapof{\cha{f}{w}}{\cha{f'}{w'}}\\
\begin{tabular}{ r l }
   & if \chapof{f}{f'} and \chapof{w}{w'}\\
or & if \chapof{f}{\cha{f'}{w'}}
\end{tabular}
\item[] \chapof{\cha{f}{w}}{f'} if \chapof{f}{f'}
\item[] \chapof{f}{f} if \scheme{operator?(f)}
\end{itemize}

%\begin{verbatim}
%chaperone-of? (chaperone f w) (chaperone f' w')
%if (chaperone-of? f f') and (chaperone-of? w w')
%;or (chaperone-of? f (chaperone f' w'))
%chaperone-of? (chaperone f w) f'
%if (chaperone-of? f f')
%chaperone-of? (e_LAMBDA,\rho) (e_LAMBDA',\rho')
%if e_LAMBDA=e_LAMBDA' and \rho=\rho'
%\end{verbatim}

\subsubsection{Operator Arity}

The arity of an operator in \chapcalc\ is a set of natural numbers each indicating a number of arguments the operator will accept.
For most operators in \chapcalc\, this arity is a singleton.
To wit, the arity of closures is simply the number of formal parameters of the $\lambda$-term.

\begin{itemize}
\item[] $\arity{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}=\{n\}$
\end{itemize}

The arity of impersonators (including chaperones) is the arity of the impersonated operator.

\begin{itemize}
\item[] $\arity{\imp{f}{w}}=\arity{f}$
\end{itemize}

The arity of each primitive procedure is straightforward with two exceptions:

\begin{itemize}
\item[] $\arity{\mathit{values}}=\{0,1,2,\dots\}$
\item[] $\arity{\mathit{raise}}=\{0,1,2,\dots\}$
\end{itemize}

\newcommand{\funarr}[2]{$\mathbf{#1}\rightarrow\mathbf{#2}$}

\begin{figure}
\label{fig:states}

\begin{tabular}{ r r l l }
\stxclass{\varsigma}{State}{$\sexp{\ks}{\sigma}{\rho}{e}$}{evaluation}\\
\stxclasscont{$\sapp{\ks}{\sigma}{f}{\vv}$}{apply}\\
\stxclasscont{$\sval{\ks}{\sigma}{\vv}$}{value}\\
\stxclasscont{$\serr{\ks}{\sigma}{b}$}{error}\\
\stxclasscont{$\simpz{\ks}{\sigma}{\ell}{f}{w}{\vv}$}{imp0}\\
\stxclasscont{$\simpo{\ks}{\sigma}{\ell}{w}{\vv}$}{imp1}\\
\stxclasscont{$\schaz{\ks}{\sigma}{\ell}{f}{w}{\vv}$}{chap0}\\
\stxclasscont{$\schao{\ks}{\sigma}{\ell}{w}{\vv}$}{chap1}\\
\stxclass{\gamma}{Frame}{\letk{\vx}{\rho}{e}}{let}\\
\stxclasscont{\impcwk{f}{\vv}}{negative guard}\\
\stxclasscont{\impcfk{w}}{impersonated procedure}\\
\stxclasscont{\impcrk{\vv}}{positive guard}\\
\stxclasscont{\chacwk{f}{\vv}}{chaperone negative}\\
\stxclasscont{\chacfk{w}}{chaperone function}\\
\stxclasscont{\chacrk{\vv}}{chaperone positive}\\
\stxclass{v}{Val}{$f\,|\,n\,|\,t$}{values}\\
\stxclass{\sigma}{Store}{\funarr{Addr}{Val}}{stores}\\
\stxclass{\rho}{Env}{\funarr{Var}{Addr}}{environments}\\
\stxclass{f,w}{Op}{\nttwo{clos}{\mathit{lam}}{\rho}}{closures}\\
\stxclasscont{\ntthr{imp}{\ell}{f}{w}}{impersonators}\\
\stxclasscont{\ntthr{chap}{\ell}{f}{w}}{chaperones}\\
\stxclasscont{$p$}{primitives}\\
\stxclass{a}{Addr}{an infinite set of addresses}{}\\
\stxclass{b}{Blame}{$\blame{\ell}\,|\,\noblame$}{blame}
\end{tabular}

\caption{Abstract machine state space for \chapcalc}
\end{figure}

\subsection{Abstract Machine States}

An $\mathbf{eval}$ state corresponds to a traditional CESK state, that is, the evaluation of an expression.
(The order of the components is reversed, however, to aid factoring out common components from among the variants.)

An $\mathbf{apply}$ state represents a point of application and dispatches on the operator type (closure, impersonator, or primitive).

A $\mathbf{value}$ state represents a point of return and dispatches on the type of the top continuation frame.
If the continuation is empty, the constituent values are the results of the program.

An $\mathbf{error}$ state represents the arisal of an error and, as the result of the program, may include blame information.

Finally, the $\mathbf{imp_0}$, $\mathbf{imp_1}$, $\mathbf{chap_0}$, and $\mathbf{chap_1}$ states each split a transition which would otherwise pop and push a frame simulaneously into a sequence of two transitions which performs each separately.
(These are for convenience for the eventual static analysis.)

\subsection{Evaluation Relation}

\begin{figure*}
\subsubsection{Application}

\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\appe{\ae_f}{\ae_1\,\dots\,\ae_n}}}{\sapp{\ks}{\sigma}{f}{\langle v_1,\dots,v_n\rangle}}
where $f=\A{\sigma}{\rho}{\ae_f}$ and $v_i=\A{\sigma}{\rho}{\ae_i}$ for $i=1,\dots,n$.

\subsubsection{The let Form}

% encounter let form
\noindent
\red{\sexp{\ks}{\sigma}{\rho}{\lete{x_1\,\dots\, x_n}{c}{e}}}{\sexp{\letk{\vx}{\rho}{e}::\ks}{\sigma}{\rho}{c}}
where $\vx=\langle x_1,\dots,x_n\rangle$.

\redtwo{\sval{\letk{\rho}{\vx}{e_1}::\ks}{\sigma}{\vv}}
       {\sexp{\ks}{\sigma'}{\rho'}{e_1}}{\mbox{if }|\vx|=|\vv|}
       {\serr{\ks}{\sigma}{\noblame}}{\mbox{if }|\vx|\ne|\vv|}
where $(\sigma',\rho')=\bind{\vx}{\vv}$

\subsubsection{if}

\redtwo{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}
       {\sexp{\ks}{\sigma}{\rho}{e_c}}{\mbox{if }v_t\ne\false}
       {\sexp{\ks}{\sigma}{\rho}{e_a}}{\mbox{if }v_t=\false}
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

\subsection{apply}

\subsubsection{closure}

\redtwo{\sapp{\ks}{\sigma}{\clo{\lame{x_1\,\dots\,x_n}{e}}{\rho}}{\vv}}
       {\sexp{\ks}{\sigma'}{\rho'}{e}}{\mbox{if }|\vx|=|\vv|}
       {\serr{\ks}{\sigma}{\blame{\ell}}}{\mbox{if }|\vx|\ne|\vv|}
where $\vx=\langle x_1,\dots,x_n\rangle$ and $(\sigma',\rho')=\bind{\vx}{\vv}$

\subsubsection{impersonator}

\red{\sapp{\ks}{\sigma}{\imp{f}{w}}{\vv}}{\sapp{\impcwk{f}{\vv}::\ks}{\sigma}{w}{\vv}}

\redtwo{\sval{\impcwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}
       {\simpz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\mbox{if }|\vv|=|\vvp|\mbox{ and } \mathrm{operator?}(w)\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\mathrm{operator?}(w)\mbox{.}}

\redtwo{\sval{\impcwk{f}{\vv}::\ks}{\sigma}{\vvp}}
       {\sapp{\ks}{\sigma}{f}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{.}}

\red{\simpz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\sapp{\impcfk{w}::\ks}{\sigma}{f}{\vvp}}

\red{\sval{\impcfk{w}::\ks}{\sigma}{\vv}}{\simpo{\ks}{\sigma}{w}{\vv}{\ell}}

\red{\simpo{\ks}{\sigma}{w}{\vv}{\ell}}{\app{\impcrk{\vv}::\ks}{\sigma}{w}{\vv}}

\redtwo{\sval{\impcrk{\vv}::\ks}{\sigma}{\vvp}}
       {\sval{\ks}{\sigma}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{ and }|\vv|+1\ne|\vvp|\mbox{.}}

\subsubsection{chaperone}

\red{\sapp{\ks}{\sigma}{\cha{f}{w}}{\vv}}{\sapp{\chacwk{f}{\vv}::\ks}{\sigma}{w}{\vv}}

\redthr{\sval{\chacwk{w}{\vv}::\ks}{\sigma}{w::\vvp}}
       {\schaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\mbox{if }|\vv|=|\vvp|\mbox{, } \mathrm{operator?}(w)\mbox{, and }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\mathrm{operator?}(w)\mbox{,}}
       {\serr{\ks}{\sigma}{\chablame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ and }\mathrm{operator?}(w)\mbox{ but not }\chapof{\vvp}{\vv}\mbox{.}}

\redthr{\sval{\chacwk{f}{\vv}::\ks}{\sigma}{\vvp}}
       {\sapp{\ks}{\sigma}{f}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{ and }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\chablame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{ and }|\vv|+1\ne|\vvp|\mbox{.}}

\red{\schaz{\ks}{\sigma}{f}{\vvp}{\ell}{w}}{\sapp{\chacfk{w}::\ks}{\sigma}{f}{\vvp}}

\red{\sval{\chacfk{w}::\ks}{\sigma}{\vv}}{\schao{\ks}{\sigma}{w}{\vv}{\ell}}

\red{\schao{\ks}{\sigma}{w}{\vv}{\ell}}{\app{\chacrk{\vv}::\ks}{\sigma}{w}{\vv}}

\redthr{\sval{\chacrk{\vv}::\ks}{\sigma}{\vvp}}
       {\sval{\ks}{\sigma}{\vvp}}{\mbox{if }|\vv|=|\vvp|\mbox{ and }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\chablame{\ell}}}{\mbox{if }|\vv|=|\vvp|\mbox{ but not }\chapof{\vvp}{\vv}\mbox{,}}
       {\serr{\ks}{\sigma}{\impblame{\ell}}}{\mbox{if }|\vv|\ne|\vvp|\mbox{.}}

\subsubsection{primitive}
\red{\sapp{\ks}{\sigma}{p}{\vv}}{\sval{\ks}{\sigma}{\delta(p,\vv)}}
if $\delta(p,\vv)=\vvp$ for some $\vvp$.

\red{\sapp{\ks}{\sigma}{p}{\vv}}{\serr{\ks}{\sigma}{\delta(p,\vv)}}
if $\delta(p,\vv)=b$ for some $b$.
\end{figure*}

\subsubsection{Application}

\subsubsection{The let Form}

\subsubsection{if}

\red{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}{\sexp{\ks}{\sigma}{\rho}{e_c}}
if $v_t\ne\false$
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

\red{\sexp{\ks}{\sigma}{\rho}{\ife{\ae_t}{e_c}{e_a}}}{\sexp{\ks}{\sigma}{\rho}{e_a}}
if $v_t=\false$
where $v_t=\A{\sigma}{\rho}{\ae_t}$.

\subsection{apply}

\emph{apply} states serve as the branching point for dispatch on the operator.

\subsubsection{closure}

The application of a closure operator proceeds straightforwardly by extending its environment and evaluating its body.

If the number of values applied is incompatible with the arity of the closure, blame is assigned dynamically.

\subsubsection{impersonator}

Impersonator application recurs on its guard and augments the context to complete the interposition when its guard returns.
The enlarged context is annotated with the label of the applied chaperone (which itself is the label of its installment site).

When the results of its wrapper are obtained, the next state is determined by the results.
If an additional result is provided and it is an operator, it is situated to intercept procedure results.

If the number of results obtained is the same as the number of arguments given, the interposition ceases and the function can be evaluated in tail position with respect to the chaperone application.

[That the number of arguments received is the same as the number given is ensured.]

\subsubsection{chaperone}

Chaperone application recurs on its guard and augments the context to complete the interposition when its guard returns.
The enlarged context is annotated with the label of the applied chaperone (which itself is the label of its installment site).

When the results of its wrapper are obtained, the next state is determined by the results.
If an additional result is provided and it is an operator, it is situated to intercept function results.

If the number of results obtained is the same as the number of arguments given, the interposition ceases and the function can be evaluated in tail position with respect to the chaperone application.

[DEFINE BLAME!]

The label of the call site is passed implicitly to each primitive.
The \scheme{imp-op} and \scheme{chap-op} primitives imbue the newly-created value with it.
Otherwise, it is used to blame the context.

\newcommand{\primone}[4]{$\delta$(#1,$\langle #2 \rangle$) &= $\langle #3 \rangle$ & #4}
\newcommand{\primtwo}[5]{$\delta$(#1,$\langle #2, #3 \rangle$) &= $\langle #4 \rangle$ & #5}
\newcommand{\primtwe}[4]{$\delta$(#1,$\langle #2, #3 \rangle$) &= $\ell$ & #4}
\newcommand{\primfou}[7]{$\delta$(#1,$\langle #2, #3, #4, #5\rangle$) &= $\langle #6 \rangle$ & #7}
\newcommand{\primfoe}[6]{$\delta$(#1,$\langle #2, #3, #4, #5\rangle$) &= $\ell$ & #6}
\begin{tabular}{ r l l }
\primtwo{\scheme|<|}{n_1}{n_2}{\true}{if $n_1 < n_2$}\\
\primtwo{\scheme|<|}{n_1}{n_2}{\false}{otherwise}\\
\primtwo{\scheme|>|}{n_1}{n_2}{\true}{if $n_1 > n_2$}\\
\primtwo{\scheme|>|}{n_1}{n_2}{\false}{otherwise}\\
\primtwo{\scheme|=|}{n_1}{n_2}{\true}{if $n_1 = n_2$}\\
\primtwo{\scheme|=|}{n_1}{n_2}{\false}{otherwise}\\
\primtwo{\scheme|+|}{n_1}{n_2}{\lceil n_1 + n_2\rceil}{}\\
\primtwo{\scheme|-|}{n_1}{n_2}{\lceil n_1 - n_2\rceil}{}\\
\primtwo{\scheme|imp-op|}{f_1}{f_2}{\imp{f_1}{f_2}}{if $\arity{f_2}\supseteq\arity{f_1}$}\\
\primtwe{\scheme|imp-op|}{f_1}{f_2}{otherwise}\\
\primtwo{\scheme|chap-op|}{f_1}{f_2}{\cha{f_1}{f_2}}{if $\arity{f_2}\supseteq\arity{f_1}$}\\
\primtwe{\scheme|chap-op|}{f_1}{f_2}{otherwise}\\
\primfou{\scheme|values|}{v_1}{v_2}{\dots}{v_n}{v_1,v_2,\dots,v_n}{}\\
\primfoe{\scheme|raise|}{v_1}{v_2}{\dots}{v_n}{}\\
\primone{\scheme|not|}{v}{\true}{if $v = \false$}\\
\primone{\scheme|not|}{v}{\false}{otherwise}\\
\primone{\scheme|operator?|}{f}{\true}{}\\
\primone{\scheme|operator?|}{v}{\false}{for $v$ not an operator}\\
\primone{\scheme|integer?|}{n}{\true}{}\\
\primone{\scheme|integer?|}{v}{\false}{for $v$ not an integer}\\
\primone{\scheme|boolean?|}{t}{\true}{}\\
\primone{\scheme|boolean?|}{v}{\false}{for $v$ not a boolean}\\
\end{tabular}

The class $n$ of variables indicate integers, the class $f$ operators, the class $t$ booleans, and the class $v$ a union of all value types.
If a primitive is applied to a different number or type of arguments than shown, its evaluation raises an error.

The \scheme{values} primitive is distinct.
It takes any number of arguments and returns that same number.
Its arity is taken to be equal to the arity of any other operator, including itself.
Thus, the installation \scheme{(chap-op values values)} is legal, if superfluous).

The reduction relation is defined as the union of each given relation.
[show it doesn't get stuck somewhere else?]
\begin{itemize}
\item If $I(p)\longrightarrow^{*}\sval{\langle\rangle}{\sigma}{\vv}$, the program evaluates to the values $\vv$ which we write as $p\Downarrow\vv$.
\item If $I(p)\longrightarrow^{*}\serr{\langle\gamma_1,\dots,\gamma_n\rangle}{\sigma}{b}$, $p$ the program results in an error with blame $b$ and we write $p\Uparrow b$.
\item If, for all $\varsigma$ such that $I(p)\longrightarrow^{*}\varsigma$, there exists $\varsigma'$ such that $\varsigma\longrightarrow\varsigma'$, then $p$ diverges and we write $p\Uparrow$.
\end{itemize}

and remember to define a program resulting in a blame error, diverging, etc.