### 2013-11-16 pldi2014/content/evaluation.tex

\section{Evaluation}

It is difficult to assess the effectiveness of our analysis since interesting programs tend to use Racket features presently unsupported.
(We discuss limitations and strategies to overcome them in Section~\ref{sec:limitations}.)

We will consider a few cases which highlight its particular effectiveness.

First, we apply it to our motivating example of the cost problem, the factorial program with ``deeply-installed'' chaperones.
This example highlights the importance of choosing a good abstract domain for base values as the coincidence of the structure of the abstract domain with the boundaries of contract domains can significantly influence the precision of the analysis.

In our running \scheme{fact} example, abstracting the integers to a singleton domain by way of the definition $\alpha(n)=\bullet$ for $\bullet$ the sole element prevents the analysis from discharging a single chaperone.
It does admit that the chaperone can be demoted, but that does not remove the space cost that motivated the need for such an analysis in the first place.
However, by abstracting integers to the domain of signs defined by
\begin{equation*}
\alpha(n)=\begin{cases}\{-\} &\mbox{if }n<0\mbox{,}\\\{0\} &\mbox{if }n=0\mbox{,}\\\{+\} &\mbox{if }n>0\mbox{,}\end{cases}
\end{equation*}
and abstracting arithmetic primitives in the obvious way, our analysis can discharge the chaperone completely (assuming that constraint information is propagated over conditionals).
This effects a 70\% decrease in time cost of this procedure.

However, this example does not assess the effectiveness of the analysis of higher-order chaperone discharge.
To do so, we perform it on a version of the chaperoned factorial program which uses Church numerals, originally used by Strickland et al.~\cite{strickland2012chaperones} to benchmark the performance impact of chaperones.

This program makes heavy use of a chaperone-installation wrapper \scheme{->} to treat contract installation like contract ascription.
Using the Racket contract definitions of \scheme{any/c} as \scheme{(LAMBDA (x) x)} and \scheme{any} as \scheme{values}, the program also makes heavy use of the custom \scheme{church/c} contract defined as \scheme{(-> (-> any/c any) (-> any/c any))}.
Church numerals are chaperoned by the wrapper \scheme{(-> church/c church/c)}.
In short, higher-order chaperones abound in this small program.

Because all chaperone installation is directed through the \scheme{->} operator, there is only one chaperone-installation site in the program.
We therefore inline its definition to allow a more granular analysis increasing the number of sites to 10.
This is ultimately unnecessary as the analysis is able to discharge every site.
For the evaluation of \scheme{(fac 6)}, chaperoning accounts for just under 90\% of the time cost.

%Because higher-order contracts often refine domains, the choice of abstraction function can have significant ramifications for the precision of the analysis.
%For instance, an analysis of the program
%\begin{schemedisplay}
%(letrec ([(fact) (values (LAMBDA (n)
%                          (let ([test0 (= n 0)])
%                            (if test0
%                                1
%                                (let ([(n-1) (- n 1)])
%                                  (let ([(rec) (fact n-1)])
%                                    (* n rec)))))))])
%  (let ([(fact) (-> fact nonnegative? positive?)])
%    (fact 5)))
%\end{schemedisplay}
%using a singleton domain [?] for integers will fail to discharge the chaperone.
%A domain which distinguishes values coincident with the edges of the contract would be able to.

%\begin{tabular}{

%\end{tabular}